---
title: "Data Scientist Skills"
author: "Marley Myrianthopoulos, Sean Amato, and Molly Siebecker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gathering and Tidying Data

```{r}
## load libraries
library(tidyr)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(scales)

data_scientist_skills_list <- read.csv("https://raw.githubusercontent.com/samato0624/DATA607_Project3/main/docs/data_scientist_skills.csv")

data_scientist_skills_list <- data_scientist_skills_list[-1,]

unique_skills_list <- data.frame(skills_list = unlist(data_scientist_skills_list))

unique_skills_list <- unique_skills_list %>%
  filter(skills_list != "") %>%
  group_by(skills_list) %>%
  mutate(count = length(skills_list)) %>%
  distinct(skills_list,count) %>%
  arrange(desc(count), skills_list)
  
rownames(unique_skills_list) <- NULL

final_skills_list <- unique_skills_list[c(1:9,12,13,17,19,22,24,27:29,32,34,36:37,39,40,43,45,47:50,52,53,58,59,61,62,64,66,71,72,76,77,79:81,83:87),1]

final_skills_list <- final_skills_list %>% arrange(skills_list)

final_skills_list$skills_list <- tolower(final_skills_list$skills_list)

final_skills_list$skills_list[38] <- " r "

kable(final_skills_list, format = "pipe", col.names = c(""), caption = "Skills for Data Scientists", align = "c")

```

```{r}

raw_data <- read.csv("https://raw.githubusercontent.com/samato0624/DATA607_Project3/main/Raw_Data.csv")

skills <- raw_data %>%
  
  # removed unnecessary characters
  mutate(Pay.Range = (str_remove_all(Pay.Range, "[$yr////, ]"))) %>%
  
  # extracted the low end of the pay range
  mutate(Pay_Min = as.numeric(str_extract(Pay.Range, "\\d+(?=-)"))) %>%
  # extracted the high end of the pay range
  mutate(Pay_Max = as.numeric(str_extract(Pay.Range, "(?<=-)[0-9]+"))) %>%
  
  # selected important columns
  select(c(2, 3, 5, 6, 7, 8))
```

# Analysis

"Most valued skills" can mean either the skills that the most people want, or the skills that people want the most. Therefore, we are going to focus our analysis on two measures: number of postings and salary. 

To analyze which skills are the most valuable by salary, we are going to focus on the minimum of the given salary range for each posting. When a skill is particularly highly valued, the minimum salary in the range would not go below a certain amount. The inverse, however, is not necessarily true. For a skill that is not particularly highly valued, we cannot assume that the maximum salary would not go above a certain amount. Skills that are less highly valued are likely to be considered baseline requirements that would appear in job postings for higher-paying as well as lower-paying jobs, making the maximum salary less meaningful for comparison than the minimum salary. 

```{r}

skills_data <- final_skills_list


for(i in 1:nrow(skills_data)) {
  skills_data$postings[i] <- as.numeric(skills %>% filter(str_detect(tolower(About.the.job),skills_data$skills_list[i])) %>% summarise(n = n()))
}

skills_data <- skills_data %>% filter(postings != 0)

for(i in 1:nrow(skills_data)) {
  skills_data$avg_min_sal[i] <- as.numeric(skills %>% drop_na(Pay_Min) %>% filter(str_detect(tolower(About.the.job),skills_data$skills_list[i])) %>% summarise(avg = mean(Pay_Min)))
}

skills_data$avg_min_sal <- dollar(skills_data$avg_min_sal)
kable(skills_data, format = "pipe", col.names = c("Skill", "Number of Postings", "Average Minimum Salary"), caption = "Skills for Data Scientists", align = "c", digits = 0)
```

Below, we filter the list of skills to include only those with at least 5 postings, arrange in descending order by number of postings, and display the frequency of each skill in a bar graph.

```{r}
## filter for five or more postings
skills_data_filtered <- skills_data %>% filter(postings >= 5)

## arrange in descending order by number of postings
skills_data_filtered$skills_list <- factor(skills_data_filtered$skills_list, levels = skills_data_filtered$skills_list[order(skills_data_filtered$postings, decreasing = FALSE)])

## create bar graph
ggplot(skills_data_filtered, aes(x = skills_list, y = postings, fill = skills_list)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "Number of Postings by Skill", x = "Skill", y = "Frequency") +
  coord_flip() 
```

Below, we order the skills list in descending order by average minimum salary, reformat the average minimum salary as a number, and display the averge minimum salary of each skill in a bar graph.

```{r}
## arrange in descending order by average minimum salary
skills_data_filtered$skills_list <- factor(skills_data_filtered$skills_list, levels = skills_data_filtered$skills_list[order(skills_data_filtered$avg_min_sal, decreasing = FALSE)])

## convert average minimum salary to number
skills_data_filtered$avg_min_sal <- parse_number(skills_data_filtered$avg_min_sal)

## create bar graph
ggplot(skills_data_filtered, aes(x = skills_list, y = avg_min_sal, fill = skills_list)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "Average Minimum Salary by Skill", x = "Skill", y = "Average Minimum Salary (dollars per year)") +
  coord_flip()
```

# Findings and Recommendations

There are only two skills that appear in the top five skills when measured both by frequency and by average minimum salary: communication and leadership; both are "soft skills." This is evidence that soft skills like communication and leadership are highly valued and should not be neglected by aspiring data scientists in favor of hard skills. 

In the context of big organizations the observed phenomena makes sense. Data scientists with technical skills and good leadership are unicorns. They drive organizations and develop the talent of aspiring data scientists. The ability to persuade people and communicate findings will determine whether or not people take your advice with regards to strategic decision making, but having the ability to guide other data scientists to do this is what brings value to the business.

We also note that Python and SQL are very frequently mentioned skills, and R ranks 12th both in number of postings and average minimum salary. Python is listed over 5 times as often as R, and corresponds to a more than $15,000 average minimum salary increase. SQL is listed almost 3 times as often as R, and corresponds to almost a $20,000 average minimum salary increase.



